---
title: "USA Well Water Data"
author: "Annette Hilton"
date: "12/01/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}

# Attach packages 

library(tidyverse)
library(kableExtra)

```


## Goals and Objectives for the project 

This project compiles existing groundwater data in the United States of America from the United States Geological Survey (USGS) across all 50 states and 4 territories (District of Columbia, Guam, Puerto Rico, Virgin Islands). 

The code automatically downloads these data from USGS sites using URLs with little user input. 

### Data sources 

#### USGS

All USGS Groundwater Well Records were obtained during the month of December, 2021. 

USGS data has been obtained through two different download portals:

- USGS Groundwater Levels REST Web Service URL Generation Tool
  - https://waterservices.usgs.gov/rest/GW-Levels-Test-Tool.html
  
- National Water Information System: Web Interface
  - https://nwis.waterdata.usgs.gov/nwis/si
  **Make sure you are on the Groundwater Data for the Nation Site Information webpage (not just groundwater), and then search by individual state (otherwise you will not get full data)
  
### Parameters included in download 

**USGS Groundwater Levels REST Web Service URL Generation Tool** 

- agency_cd     Agency code
- site_no       USGS site number
- site_tp_cd    Site type code
- lev_dt        Date level measured
- lev_tm        Time level measured
- lev_tz_cd     Time datum
- lev_va        Water-level value in feet below land surface
- sl_lev_va     Water-level value in feet above specific vertical datum
- sl_datum_cd   Referenced vertical datum
- lev_status_cd Status
- lev_agency_cd Measuring agency
- lev_dt_acy_cd Water-level date-time accuracy
- lev_acy_cd    Water-level accuracy
- lev_src_cd    Source of measurement
- lev_meth_cd   Method of measurement
- lev_age_cd    Water-level approval status 

**USGS National Water Information System: Web Interface** 

- agency_cd       -- Agency
- site_no         -- Site identification number
- station_nm      -- Site name
- dec_lat_va      -- Decimal latitude
- dec_long_va     -- Decimal longitude
- coord_acy_cd    -- Latitude-longitude accuracy
- dec_coord_datum_cd -- Decimal Latitude-longitude datum
- state_cd        -- State code
- alt_va          -- Altitude of Gage/land surface
- alt_acy_va      -- Altitude accuracy
- alt_datum_cd    -- Altitude datum
- well_depth_va   -- Well depth
- hole_depth_va   -- Hole depth


#### How to download data (generate URL) or change parameters 

##### **USGS Groundwater Levels REST Web Service URL Generation Tool** 

Required Arguments Section 

- Major Filters: "State or Territory" 
- State or Territory: Select state of interest 

Optional Arguments Section 

- Output Format: "USGS RDB (tab-delimited)" 
- Date Range: "Return all values within an absolute date range (start and end dates)" 
- Start Date: "1800-01-01" 
- Groundwater Site Types: "Well" 

Generated URL Section 

- Click "Generate the URL" 
- Copy the URL (or click "Run the Generated URL" and download the data)

##### **USGS National Water Information System: Web Interface** 

- Select Geographic Area: Select state of interest
- Site Attribute: "Site type" 
- Click "Submit" 

(Next page)

- Site type: "Well" 
- Choose Output format: "Site-description information displayed in tab separated format 
- Select Fields to include in site-description output: 
  - Agency
  - Site identification number
  - Site name 
  - Decimal latitude 
  - Decimal longitude 
  - Altitude of Gage/land surface
  - Well depth 
  - Hole depth 
- Click "Submit" 
- Copy the URL or save the results as a text file 

