---
title: "USA Well Water Data"
author: "Annette Hilton"
date: "10/31/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}

# Attach packages 

library(tidyverse)
```


## Goals and Objectives for the project 

Here we attempt to compile existing groundwater data in the United States of America across all 50 states. 
We will begin by obtaining and compiling all available USGS groundwater data. We will then map well locations to identify where data exists and where data is scarce. 
We will prioritize states and locations with little data to investigate if other groundwater data exists and if so, how to obtain it. 
We will process and clean all of the data to ensure that it is in a consistent and computationally friendly format. 
After we have obtained as much groundwater data as reasonably possible and converted it into a consistent format, we hope to have a large dataset in which we can begin to explore trends in groundwater. 

## Specific parameters 

We are seeking the following parameters: 

- Location of well (latitude, longitude)
- Date of measurement 
- Well name/identifier 
- Well water depth 
- Well screen depth 
- Elevation of well 

## Constraints of data 

We are specifically seeking both historical and continuous groundwater data to provide us with information about how the water level changed over time. 

For each data type we plan to constrain our compilation by the following factors: 

**Historical data**

- Data from historical wells with only one measurement (well depth when drilled) will be excluded 

**Continuous data** 

- Any continuous data measurements that have more than one measurement per day will be averaged to obtain the mean well water depth for the entire day 
